---
title: "In-class Exercise 5"
author: "Nguyen Nguyen Ha (Summer)"
date: "May 17, 2025"
date-modified:  "May 17, 2025"
format:
  html:
    css: styles.css
    toc: true
execute: 
  warning: false
  freeze: true
---

# 1. Getting Started
## 1.1. Loading Packages 

In the code chunk below, `p_load()` of pacman package is used to load the R packages into R environment
```{r}
pacman::p_load(tidyverse, jsonlite,
               SmartEDA, tidygraph,
               ggraph)
```

## 1.2. Importing JSON 

In the code chunk, `fromJSON()` of **jsonlite** package is used to import **MC1_graph.json** file into R and save the output subject

```{r}
kg <- fromJSON("data/MC1_graph.json")
```

## 1.3. Inspecting structure
```{r}
str(kg, max.level = 1)
```

## 1.4. Extracting and inspecting
```{r}
nodes_tbl <- as_tibble(kg$nodes)
edges_tbl <- as_tibble(kg$links)
```

# 2. Initial EDA

## 2.1. Edge Type
```{r}
ggplot(data = edges_tbl, 
       aes(y = `Edge Type`, fill = `Edge Type`)) +
  geom_bar()
```

## 2.2. Node Type
```{r}
ggplot(data = nodes_tbl, 
       aes(y = `Node Type`, fill = `Node Type`)) +
  geom_bar()
```

# 3. Creating Knowledge Graph
## 3.1. Step 1: Mapping from node ID to row Index
```{r}
id_map <- tibble(id=nodes_tbl$id,
                 index = seq_len(nrow(nodes_tbl)))
```
This ensures each id from your node list is mapped to the correct row number. 

## 3.2. Step 2: Map source and target IDs to row indices 
```{r}
edges_tbl <- edges_tbl %>%
  left_join(id_map, by = c("source" = "id")) %>%

  rename(from = index) %>%
  left_join(id_map, by = c("target" = "id")) %>%

  rename(to = index)
```

## 3.3. Step 3: Filter out any unmatched (invalid) edges
```{r}
edges_tbl <- edges_tbl %>%
  filter(!is.na(from), !is.na(to))
```

## 3.4. Step 4: Creating the graph

Lastly, `tbl_graph` is used to create tidygraph's graph object by using the code junk below.
```{r}
graph <- tbl_graph(nodes = nodes_tbl, 
                   edges = edges_tbl,
                   directed = kg$directed)
```

# 4. Visualizing the knowledge graph
```{r}
set.seed(1234)
```

## 4.1. Visualizing the whole graph
```{r, eval=FALSE}
ggraph(graph, layout ="fr") + 
  geom_edge_link(aplpha = 0.3, colour = "gray") +
  geom_node_point(aes(color = `Node Type`), size = 4) +
  geom_node_text(aes(label = name), repel = TRUE, size = 2.5) +
  theme_void()
```

## 4.2. Visualizing the sub-graph

In this section, we are interested to create a sub-graph based on `MemberOf` value in `Edge_Type` column of the `edges` column of the `edge` data frame.

### 4.2.1. Step 1: Filter edges to only `MemberOf'
```{r}
graph_memberof <- graph %>%
  activate(edges) %>% 
  filter(`Edge Type` == "MemberOf")
```

### 4.2.2. Step 2: Extract only connected nodes (i.e., used in these edges)
```{r}
used_node_indices <- graph_memberof %>%
  activate(edges) %>%
  as_tibble() %>%
  select(from, to) %>%
  unlist() %>%
  unique()
```

### 4.2.3. Step 3: Keep only those nodes
```{r}
graph_memberof <- graph_memberof %>%
  activate(nodes) %>%
  mutate(row_id = row_number()) %>%
  filter(row_id %in% used_node_indices) %>%
  select(-row_id)
```

### 4.2.4. Step 4: Plot the sub-graph
```{r}
ggraph(graph_memberof, layout = "fr") + 
  geom_edge_link(aplpha = 0.5, colour = "gray") +
  geom_node_point(aes(color = `Node Type`), size = 1) +
  geom_node_text(aes(label = name), repel = TRUE, size = 2.5) +
  theme_void()
```
